Stochastic Variance-Reduced Policy Gradient

### 2018-02-05 svrpg.tex
Added abstract and checked intro

\begin{abstract}
Why we need variance-reduced gradient in RL?
\end{abstract}

@@

In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$. The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution. When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM). Even in this case, stochastic optimization is often used for reasons of efficiency. The idea of stochastic gradient ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction. At the level of the single iteration, this can be much less expensive than taking into account all the data. However, the sub-sampling of data samples is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of addressing this trade-off, with significant results both in theory and practice. Although designed explicitly for ERM, this algorithms address a problem that affects more general machine learning problems. 

@@

Among RL approaches, policy gradient \cite{sutton2000policy} is the one that bears the closest similarity to SL solutions. The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient ascent. Compared to other applications of SGD, the cost of collecting samples can be very high since it requires to interact with the environment. This makes SVRG-like methods potentially much more efficient than, \eg batch learning.

@@

%Paper structure
After providing a background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address the aforementioned difficulties to design SVRPG, a variant of SVRG for the policy gradient framework. In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL. In Section \ref{sec:stopping} we suggests how to set the hyperparameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm. Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL benchmark tasks.

\section{Preliminaries}\label{sec:pre}
We start providing the essential background on policy gradient methods for Reinforcement Learning and stochastic variance-reduced gradient methods for finite-sum optimization.
%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

@@

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
%of $J(\vtheta)$ \wrt the policy parameters 
is \cite{sutton2000policy, Peters2008reinf}:
\begin{align} \label{E:policygradient}
        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align}
Notice that the distribution defining the gradient is induced by the current policy. This aspect introduces a nonstationarity in the sampling process. Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
In this setting, stochastic gradient ascent is typically employed.
At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$, \ie
\begin{align} \label{E:policygradient.estimate}
        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
\end{align}
The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
For example, REINFORCE is such that $g(\tau|\vtheta) = \nabla \log\pi(\tau|\vtheta) \mathcal{R}(\tau)$. 
Although REINFORCE definition is simpler, G(PO)MDP is often preferred due to its lower variance.
We refer the reader to App.~\ref{A:gradient_estimators} for details.

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
Recall that compared to SL, the samples are not fixed in advance but need to be collected at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
Due to this issues, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following properties:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.\todopir{this should be shown in appendix}
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. {\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.} 
Finally, the unbiasedness of Eq.~\ref{E:svrg.gradient.correction} is broken by the non-stationarity of the optimization problem.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\ref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
Our update is still fundamentally on-policy, since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
        \right].
\end{align}
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
\begin{lemma}
Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:
\[
	\Var\left[\gradBlack{\vtheta^*}\right] = 
    \Var\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
\]
\end{lemma}
For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.

@@

\textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
In practice, due to the lower variance G(PO)MDP is preferred to REINFORCE.
For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
\todopir{sketch of proof for GPOMDP in appendix}

@@

\subsection{Normalized Importance Sampling}\label{sec:prac}
\todopir{TODO: This section should be generic for normalized importance sampling trying to explain when and where are the advantages of doing this. Explicit formulation can be postponed in appendix.}
\todomat{Missing opening to section}
We introduce a term with importance weights to guarantee proper sampling of trajectories.
This term's variance is subjected to the variance of importance weights. 
There are some methods to reduce this variance.

As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance. The off-policy gradient estimator becomes:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
\end{align*} 
where:
\begin{align*}
\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
 }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\begin{align*}
\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
\section{Related Work}

@@

The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline provided $\EVV[a \sim \pol]{b(s,a)} = 0$, for any $s\in\Sspace$.
The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}

@@

			+\frac{1}{B}\sum_{i=0}^{B-1} g(\tau_i\vert\vtheta_t^{s+1}) 

@@

			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

			g(\tau_i\vert\vtheta_t^{s+1}) -
			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)

@@

			g(\tau_i\vert\vtheta_t^{s+1}) -
				\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 

@@

			g(\tau_i\vert\vtheta_t^{s+1}) -
			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}^2} \label{eq:2}\\

@@

			g(\tau_i\vert\vtheta_t^{s+1}) -
			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
			-g(\tau_i\vert\tilde{\vtheta}^s)
			+g(\tau_i\vert\tilde{\vtheta}^s) 
			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

			\Ets{\norm[]{g(\tau_i\vert\vtheta_t^{s+1})
			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

			\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

>>>>

\begin{abstract}
        We present and analyse stochastic variance-reduced gradient (SVRG) in the context of Markov Decision Processes (MDPs).
        We first show how to account for I) non-concativity of the objective function; II) approximations in the full gradient computation; and III) non-stationarity of the sampling process.
        These characteristics make the application of SVRG to MDPs challenging.
        We propose SVRPG, a stochastic variance-reduced policy gradient algorithm that leverages on importance weight to preserve the unbiasedness of the gradient estimate.
        We further analyse its convergence guarantees under common assumptions on the MDP.
        Finally, we investigate practical variants of SVRPG and we empirically evaluate them on continuous MDPs.
\end{abstract}

@@

In a typical SL setting, a performance function $J(\vtheta)$ has to be optimized \wrt to model parameters $\vtheta$. The set of data that are available for training is often a subset of all the cases of interest, which may even be infinite, leading to optimization of finite sums to approximate expected performance over an unknown data distribution. When generalization to the complete dataset is not taken into consideration we talk about Empirical Risk Minization (ERM). Even in this case, stochastic optimization is often used for reasons of efficiency. The idea of stochastic gradient (SG) ascent \cite{nesterov2013introductory} is to iteratively focus on a random subset of the available data to obtain an approximate improvement direction. At the level of the single iteration, this can be much less expensive than taking into account all the data. However, the sub-sampling of data is a source of variance that can potentially compromise convergence, so that per-iteration efficiency and convergence rate must be traded off with proper handling of meta-parameters.
Variance-reduced gradient algorithms such as SAG \cite{roux2012stochastic}, SVRG \cite{johnson2013accelerating} and SAGA \cite{defazio2014saga} offer better ways of addressing this trade-off, with significant results both in theory and practice. Although designed explicitly for ERM, these algorithms address a problem that affects more general machine learning problems. 

@@

Among RL approaches, policy gradient \cite{sutton2000policy} is the one that bears the closest similarity to SL solutions. The fundamental principle of this methods is to optimize a parametric policy by means of stochastic gradient ascent. Compared to other applications of SG, the cost of collecting samples can be very high since it requires to interact with the environment. This makes SVRG-like methods potentially much more efficient than, \eg batch learning.

@@

%Paper structure
After providing a background on policy gradient and SVRG in Section \ref{sec:pre}, in Section \ref{sec:alg} we address the aforementioned difficulties to design SVRPG, a variant of SVRG for the policy gradient framework. In Section \ref{sec:conv} we provide convergence guarantees for our algorithm which also offers insights on the challenges of variance reduction in RL. In Section \ref{sec:stopping} we suggests how to set the meta-parameters of SVRPG, while in Section \ref{sec:prac} we discuss some practical variants of the algorithm. Finally, in Section \ref{sec:exp} we empirically evaluate the performance of our method on popular continuous RL tasks.

\section{Preliminaries}\label{sec:pre}
We start providing the essential background on policy gradient methods and stochastic variance-reduced gradient methods for finite-sum optimization.
%In the next section we will show how these methods can be combined to solve Reinforcement Learning tasks.

@@

Policy gradient methods restrict the search for the performance-maximizing policy over a class of parametrized policies $\Pi_{\vtheta}=\{\pol: \vtheta \in \realspace^d\}$, with the only constraint that $\pol$ is differentiable \wrt $\vtheta$. We denote the performance of a parametric policy with $J(\vtheta)$ for brevity, while in some occasions we replace $\pol(\tau)$ with $\pi(\tau|\vtheta)$ for the sake of readability.
The search for a locally optimal policy is performed by means of gradient ascent, where the policy gradient
%of $J(\vtheta)$ \wrt the policy parameters 
is \cite{sutton2000policy, Peters2008reinf}:
\begin{align} \label{E:policygradient}
        \gradJ{\vtheta} = \EVV[\tau \sim \pol]{\score{\vtheta}{\tau}\Reward(\tau)}
        %\int_{\Tspace}\pol(\tau)\score{\vtheta}{\tau}\Reward(\tau)\de \tau.
\end{align}
Notice that the distribution defining the gradient is induced by the current policy. This aspect introduces a nonstationarity in the sampling process. Since the underlying distribution changes over time, it is necessary to resample at each update or use weighting techniques such as importance sampling.
Here, we consider the \emph{online learning scenario}, where trajectories are sampled by interacting with the environment at each policy change. 
In this setting, stochastic gradient ascent is typically employed.
At each iteration $k >0$, a batch $\mathcal{D}_N^k = \{\tau_i\}_{i=0}^N$ of $N>0$ trajectories is collected using policy $\pi_{\vtheta_k}$.
The policy is then updated as $\vtheta_{k+1}  = \vtheta_k + \alpha\gradApp{\vtheta_k}{N}$, where $\alpha$ is a step size and $\gradApp{\vtheta}{N}$ is an estimate of~\eqref{E:policygradient} using $\mathcal{D}_N^k$, \ie
\begin{align} \label{E:policygradient.estimate}
        \gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N} g(\tau_i|\vtheta), \quad \tau_i \in \mathcal{D}_N^k.
\end{align}
The most widespread gradient estimators are REINFORCE~\citep{williams1992simple} and G(PO)MDP~\citep{baxter2001infinite}.
Although REINFORCE definition is simpler, G(PO)MDP is often preferred due to its lower variance.
We refer the reader to App.~\ref{A:gradient_estimators} for details.

@@

\section{SVRG in Reinforcement Learning}\label{sec:alg}
In online RL problems the usual approach is to tune the batch size of SG in order to find the optimal trade-off between variance and speed.
Recall that compared to SL, the samples are not fixed in advance but need to be collected at each policy change.
Since this operation may be costly, we would like to minimize the number of interactions with the environment.
Due to this issues, we would like to apply SVRG to RL problems in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence.
However, a direct application of SVRG to RL is not possible due to the following properties:
% In on-line policy gradient problems, SG is not really a choice, but is dictated by the necessity of interacting with an unknown environment.\todopir{I do not understand the meaning of this sentence. I would say that SG is not a choice by referring to papers that uses batches}
% We would like to apply SVRG to the policy gradient framework in order to limit the variance introduced by sampling trajectories, which would ultimately lead to faster convergence. In this scenario, the function $f(\vtheta)$ to optimize is $J(\vtheta)$ and the (implicit) dataset $\mathcal{D}$ is the set of all possible trajectories $\mathcal{T}$. Compared to the typical finite sum optimization scenario, we have the following additional challenges:
\begin{description}
        \item[Non-concavity:] the objective function $J(\vtheta)$ is typically non-concave.
        \item[Infinite dataset:] the RL optimization cannot be expressed as a finite-sum problem. The actual dataset is infinite and coincide with $\pi_{\vtheta}(\tau)$.
%    \item[Approximation:] the dataset $\mathcal{T}$ can be infinite, but we can only sample a finite number of trajectories;
        \item[Non-stationarity:] the distribution of the samples changes over time. In particular, the value of the parameter $\vtheta$ influences the sampling process.
\end{description}
To deal with non-concavity, we require $J(\vtheta)$ to be $L$-smooth, which is a reasonable assumption for common policy classes such as truncated Gaussian and softmax~\citep[\eg][]{Furmston2012unifying,pirotta2015lipschitz}, see App.~\ref{app:gauss} for the derivation.
Because of the infinite dataset, we can only rely on an estimate of the full gradient.
\citet{harikandeh2015stopwasting} analysed this scenario under the assumptions of $z$ being \textcolor{red}{concave}, showing that SVRG is robust to inexact computation of the full gradient. In particular, it is still possible to recover the original convergence rate if the error decreases at an appropriate rate. {\color{red}Later on we will show how the estimation accuracy impacts on the convergence results with non-concave objective.} 
Finally, the unbiasedness of Eq.~\ref{E:svrg.gradient.correction} is broken by the non-stationarity of the optimization problem.
To overcome this limitation we employ importance weighting~\citep[\eg][]{rubinstein1981simulation,precup2000eligibility} to correct the distribution shift.
%Finally, we employ importance weighting~\citep[\eg]{rubinstein1981simulation,precup2000eligibility} to guarantee proper sampling of trajectories.

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\ref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{\pi(\tau|\wt{\vtheta})}{\pi(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$ and the update is a FG step.
Our update is still fundamentally on-policy, since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
        \right].
\end{align}
We want just to underlying that $\EVV[\tau_i \sim \pi_{\vtheta_t}]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$~\citep[\eg][]{jurvcivcek2012reinforcement}.\footnote{The reader can refer to App.~\ref{} for off-policy variants of REINFORCE and G(PO)MDP.}
This property will be used to prove Lem.~\ref{L:svrpg.properties}.
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see App.~\ref{app:proofs}):
\begin{lemma}\label{L:svrpg.properties}
Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:\footnote{
For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
\[
	\Var\left[\gradBlack{\vtheta^*}\right] = 
    \Var\left[\wh{\nabla}_N^{\textsc{RF}}J(\vtheta^*)\right].
\]
\end{lemma}

@@

\textbf{G(PO)MDP.} A similar result to the one stated in Thm~\ref{theo:convergence} can be derived for the G(PO)MDP estimator. A sketch of the proof is reported in~\ref{app:gpomdpproof}.
In practice, due to the lower variance G(PO)MDP is preferred to REINFORCE.
For this reason we will use G(PO)MDP for the experiments (see Sec.~\ref{sec:exp}).
\todopir{sketch of proof for GPOMDP in appendix, probably can be unified with the one of REINFORCE}

@@

\subsection{Normalized Importance Sampling}\label{sec:prac}
\todopir{TODO: This section should be generic for normalized importance sampling trying to explain when and where are the advantages of doing this. Explicit formulation can be postponed in appendix.}
We introduce a term with importance weights to guarantee proper sampling of trajectories.
This term's variance is subjected to the variance of importance weights. 
There are some methods to reduce this variance.

As reported in Section \ref{subsec:PolicyGradient} we use G(PO)MDP gradient estimator~\cite{baxter2001infinite}, which gives us the same guarantees we have for the Reinforce Estimator.
One variant in off-policy G(PO)MDP is to use the weighted per-decision importance sampling estimator \cite{precup2000eligibility}, allowing a reduction of the gradient estimator variance. The off-policy gradient estimator becomes:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{\Omega}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_k^n\vert s_k^n}\right)\left(\gamma^h r_h^n \omega(z_{0:h})\right).
\end{align*} 
where:
\begin{align*}
\omega(z_{0:h}) = \prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
 }{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\begin{align*}
\Omega = \sum_{n=1}^{N}\sum_{h=0}^{H-1}\gamma^h\prod_{k=0}^{h} \frac{\pi_{\vtheta}({a_k^n\vert s_k^n})
}{\pi_{\vtheta^{'}}({a_k^n\vert s_k^n})}
\end{align*}
\todomat{why should we use it? What advantages do we expect? At what cost? Pointer to experiments}
\section{Related Work}
\todopir{I think that this section is worth here}

@@

\subsection{On-policy}
The REINFORCE gradient estimator~\citep{williams1992simple} provides a simple, unbiased way of estimating the gradient:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\left(\sum_{h=0}^{H-1}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\sum_{h=0}^{H-1}\gamma^h r_h^n - b\right),
\end{align*}
where subscripts denote the time step, superscripts denote the trajectory, $r_h$ is the reward actually collected at time $h$ and b can be any baseline provided $\EVV[a \sim \pol]{b(s,a)} = 0$, for any $s\in\Sspace$.
The G(PO)MDP gradient estimator~\cite{baxter2001infinite} is a refinement of REINFORCE which is subject to less variance \cite{zhao2011analysis} while preserving the unbiasedness:
\begin{align*}
\gradApp{\vtheta}{N} = \frac{1}{N}\sum_{n=1}^{N}\sum_{h=0}^{H-1}\left(\sum_{k=0}^{h}\score{\vtheta}{a_h^n\vert s_h^n}\right)\left(\gamma^h r_h^n - b\right).
\end{align*}

\subsection{Off-Policy}
\todopir{Make ref to Luca's thesis and describe in brief the derivations}

@@

			+\frac{1}{B}\sum_{i=0}^{B-1} \score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) 

@@

			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\

@@

			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)

@@

			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
				\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}\right\|^2\right] 

@@

			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)\right)}^2} \label{eq:2}\\

@@

			\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i) -
			\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\

@@

			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)
			+\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
			-\omega(\tau_i)\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\

@@

			\Ets{\norm[]{\score{\vtheta_t^{s+1}}{\tau_i}R(\tau_i)
			-\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i)}^2} \nonumber\\

@@

			\Ets{\norm[]{\score{\tilde{\vtheta}^s}{\tau_i}R(\tau_i) 
